Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)

### 2019-03-01 ms.Rmd
Finish running all analyses and output to ms as drake plan

# Load packages
library(conflicted)
library(tidyverse)
library(drake)
library(here)
library(cowplot)
library(ggplot2)
library(scales)
library(here)
library(patchwork)
library(lubridate)
library(gt)
library(ggridges)

# Resolve conflicts
col_factor <- conflict_prefer("col_factor", "readr")
discard    <- conflict_prefer("discard", "purrr")
expand     <- conflict_prefer("expand", "tidyr")
filter     <- conflict_prefer("filter", "dplyr")
gather     <- conflict_prefer("gather", "tidyr")
ggsave     <- conflict_prefer("ggsave", "ggplot2")
here       <- conflict_prefer("here", "here")
intersect  <- conflict_prefer("intersect", "dplyr")
lag        <- conflict_prefer("lag", "dplyr")
Position   <- conflict_prefer("Position", "ggplot2")
setdiff    <- conflict_prefer("setdiff", "dplyr")
union      <- conflict_prefer("union", "dplyr")

source(here("code/functions.R"))

#' Set order of microclimate variables for printing.
order_vars <- c("par_total", "rh_min", "temp_mean")

#' Select microclimate variables of interest for analysis.
selected_vars <- c("rh_min", "par_total", "temp_mean")

# Load objects from drake workflow into memory
loadd(combined_morph)
loadd(daily_microclimate)
loadd(t_test_results)
loadd(paired_microclimate)
loadd(nested_paired_microclimate)
```

@@

```{r fig-5, fig.width=7, fig.height=8}
#' Extract p-values for plotting, and put into data frame
#' with xvals specifying season and yvals specifying height 
#' to plot asterisks.
p_vals <-
  t_test_results %>%
  mutate(pval = case_when(
    pval < 0.0001 ~ "***",
    pval < 0.001 ~ "**",
    pval < 0.05 ~ "*",
    TRUE ~ ""
  )) %>%
  select(season, var, pval) %>%
  spread(var, pval) %>%
  rename(
    par_total_pval = par_total,
    rh_min_pval = rh_min,
    temp_mean_pval = temp_mean
  )

# Simple function to calculate maximum y-position for plotting asterisks.
# Should be just above the mean + error bar height (1 sd).
maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}

pval_plot_data <- 
  paired_microclimate %>%
  select(-date, -month) %>%
  group_by(season, site) %>%
  summarize_all(
    maxy
  ) %>%
  group_by(season) %>%
  summarize_all(
    max
  ) %>%
  select(-site) %>%
  left_join(p_vals)

#' Format data for plotting: need columns to be mean and sd, by site and season.
paired_means <- 
  paired_microclimate %>%
  select(-date) %>%
  group_by(site, season) %>%
  summarize_all(
    list(~mean(., na.rm = TRUE),
         ~sd(., na.rm = TRUE))
  ) %>%
  mutate(
    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
  )

#' Make plot including t-test on three microclimatic variables.
#' 
#' The final plot includes multiple subplots. Each is a ridge plot showing 
#' the distribution of one of the variables between the two sites, with the
#' results of a paired t-test showing if they are different or not.
#' 
#' Need to manually set x ranges to keep x-axis scale the same across plots
x_ranges <- tibble(var = selected_vars) %>%
  mutate(
    min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
    max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
  )

# light blue / light green colorblind-friendly from colorbrewer
# paired_colors <- c("#a6cee3", "#b2df8a")
# OR dark blue / dark green colorblind-friendly from colorbrewer
# paired_colors <- c("#1f78b4", "#33a02c")
# OR dark grey / light grey for bw
paired_colors <- c("grey55", "grey90")

@@

#' Make ridge plots by mapping across nested datasets.
plot_results <-
  nested_paired_microclimate %>%
  right_join(x_ranges) %>%
  mutate(min = min - 0.05 * min,
         max = max + 0.05 * max,
         plot = pmap(list(var, data, x_min = min, x_max = max), 
                     plot_ridges, 
                     y_var = "site"),
         pval = map2(var, data, run_t_test),
         pval = case_when(
           pval < 0.0001 ~ "***",
           pval < 0.001 ~ "**",
           pval < 0.05 ~ "*",
           TRUE ~ ""
         ),
         plot = map2(
           plot, pval, 
           ~ .x + annotate("text", Inf, Inf, 
                           label = .y, hjust = 1, vjust = 1))

         # If needed, add annotation for season to make sure subplots
         # are positioned correctly

         # ,plot = map2(
         #   plot, season,
         #   ~ .x + annotate("text", -Inf, Inf,
         #                   label = .y, hjust = -1, vjust = 1)),
         # season = as_factor(season),
         # season = fct_relevel(season, "winter", "spring", "summer", "fall")

  ) %>% 
  arrange(season, var)

subplots <- plot_results$plot

#' Manually some subplots to remove redundant labels, etc.
subplots[1:9] <- map(subplots[1:9], 
                     ~ . + theme(axis.title.x = element_blank(),
                                 axis.text.x = element_blank() ))

subplots[1:12] <- map(subplots[1:12], 
                      ~ . + scale_fill_manual(
                        values = paired_colors,
                        name="Site",
                        breaks=c("okutama", "uratakao"),
                        labels=c("Okutama", "Uratakao"))
)

subplots[[10]] <- subplots[[10]] +
  labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )

subplots[[11]] <- subplots[[11]] +
  labs(x = expression("Min. Rel. Humidity (%)"))

subplots[[12]] <- subplots[[12]] +
  labs(x = expression("Mean Temp. (°C)"))

#' Add titles.
subplots <- map(subplots, ~ . + labs(title = "") )

subplots[[2]] <- subplots[[2]] +
  labs(title = "Winter\n")

subplots[[5]] <- subplots[[5]] +
  labs(title = "Spring\n")

subplots[[8]] <- subplots[[8]] +
  labs(title = "Summer\n")

subplots[[11]] <- subplots[[11]] +
  labs(title = "Fall\n")

#' Extract legend
legend <- paired_microclimate %>%
  ggplot(aes(x = temp_mean, y = site, fill = site)) +
  geom_density_ridges() +
  theme(legend.position = "bottom") +
  scale_fill_manual(
    values = paired_colors,
    name="Site",
    breaks=c("okutama", "uratakao"),
    labels=c("Okutama", "Uratakao"))

legend <- get_legend(legend)

#' Combine and write out.
subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
  subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
  subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
  subplots[[10]] + subplots[[11]] + subplots[[12]] +
  plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
```

>>>>

# Set order of microclimate variables for printing.
order_vars <- c("par_total", "rh_min", "temp_mean")

# Load objects from drake workflow into memory
loadd(selected_vars)
loadd(combined_morph)
loadd(daily_microclimate)
loadd(t_test_results)
loadd(paired_microclimate)
loadd(nested_paired_microclimate)
loadd(combined_monthly_morph)
loadd(linear_model_results)
loadd(gameto_size)
```

@@

```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8}
# #' Extract p-values for plotting, and put into data frame
# #' with xvals specifying season and yvals specifying height 
# #' to plot asterisks.
# p_vals <-
#   t_test_results %>%
#   mutate(pval = case_when(
#     pval < 0.0001 ~ "***",
#     pval < 0.001 ~ "**",
#     pval < 0.05 ~ "*",
#     TRUE ~ ""
#   )) %>%
#   select(season, var, pval) %>%
#   spread(var, pval) %>%
#   rename(
#     par_total_pval = par_total,
#     rh_min_pval = rh_min,
#     temp_mean_pval = temp_mean
#   )
# 
# # Simple function to calculate maximum y-position for plotting asterisks.
# # Should be just above the mean + error bar height (1 sd).
# maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}
# 
# pval_plot_data <- 
#   paired_microclimate %>%
#   select(-date, -month) %>%
#   group_by(season, site) %>%
#   summarize_all(
#     maxy
#   ) %>%
#   group_by(season) %>%
#   summarize_all(
#     max
#   ) %>%
#   select(-site) %>%
#   left_join(p_vals)
# 
# #' Format data for plotting: need columns to be mean and sd, by site and season.
# paired_means <- 
#   paired_microclimate %>%
#   select(-date) %>%
#   group_by(site, season) %>%
#   summarize_all(
#     list(~mean(., na.rm = TRUE),
#          ~sd(., na.rm = TRUE))
#   ) %>%
#   mutate(
#     season = factor(season, levels = c("winter", "spring", "summer", "fall"))
#   )
# 
# #' Make plot including t-test on three microclimatic variables.
# #' 
# #' The final plot includes multiple subplots. Each is a ridge plot showing 
# #' the distribution of one of the variables between the two sites, with the
# #' results of a paired t-test showing if they are different or not.
# #' 
# #' Need to manually set x ranges to keep x-axis scale the same across plots
# x_ranges <- tibble(var = selected_vars) %>%
#   mutate(
#     min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
#     max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
#   )
# 
# # light blue / light green colorblind-friendly from colorbrewer
# # paired_colors <- c("#a6cee3", "#b2df8a")
# # OR dark blue / dark green colorblind-friendly from colorbrewer
# # paired_colors <- c("#1f78b4", "#33a02c")
# # OR dark grey / light grey for bw
# paired_colors <- c("grey55", "grey90")
# 
# subplots <- list()
# 
# #' Make ridge plots by mapping across nested datasets.
# plot_results <-
#   nested_paired_microclimate %>%
#   right_join(x_ranges) %>%
#   mutate(min = min - 0.05 * min,
#          max = max + 0.05 * max,
#          plot = pmap(list(var, data, x_min = min, x_max = max), 
#                      plot_ridges, 
#                      y_var = "site"),
#          pval = map2(var, data, run_t_test),
#          pval = case_when(
#            pval < 0.0001 ~ "***",
#            pval < 0.001 ~ "**",
#            pval < 0.05 ~ "*",
#            TRUE ~ ""
#          ),
#          plot = map2(
#            plot, pval, 
#            ~ .x + annotate("text", Inf, Inf, 
#                            label = .y, hjust = 1, vjust = 1))
#          
#          # If needed, add annotation for season to make sure subplots
#          # are positioned correctly
#          
#          # ,plot = map2(
#          #   plot, season,
#          #   ~ .x + annotate("text", -Inf, Inf,
#          #                   label = .y, hjust = -1, vjust = 1)),
#          # season = as_factor(season),
#          # season = fct_relevel(season, "winter", "spring", "summer", "fall")
#          
#   ) %>% 
#   arrange(season, var)
# 
# subplots <- plot_results$plot
# 
# #' Manually some subplots to remove redundant labels, etc.
# subplots[1:9] <- map(subplots[1:9], 
#                      ~ . + theme(axis.title.x = element_blank(),
#                                  axis.text.x = element_blank() ))
# 
# 
# subplots[1:12] <- map(subplots[1:12], 
#                       ~ . + scale_fill_manual(
#                         values = paired_colors,
#                         name="Site",
#                         breaks=c("okutama", "uratakao"),
#                         labels=c("Okutama", "Uratakao"))
# )
# 
# subplots[[10]] <- subplots[[10]] +
#   labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )
# 
# subplots[[11]] <- subplots[[11]] +
#   labs(x = expression("Min. Rel. Humidity (%)"))
# 
# subplots[[12]] <- subplots[[12]] +
#   labs(x = expression("Mean Temp. (°C)"))
# 
# #' Add titles.
# subplots <- map(subplots, ~ . + labs(title = "") )
# 
# subplots[[2]] <- subplots[[2]] +
#   labs(title = "Winter\n")
# 
# subplots[[5]] <- subplots[[5]] +
#   labs(title = "Spring\n")
# 
# subplots[[8]] <- subplots[[8]] +
#   labs(title = "Summer\n")
# 
# subplots[[11]] <- subplots[[11]] +
#   labs(title = "Fall\n")
# 
# #' Extract legend
# legend <- paired_microclimate %>%
#   ggplot(aes(x = temp_mean, y = site, fill = site)) +
#   geom_density_ridges() +
#   theme(legend.position = "bottom") +
#   scale_fill_manual(
#     values = paired_colors,
#     name="Site",
#     breaks=c("okutama", "uratakao"),
#     labels=c("Okutama", "Uratakao"))
# 
# legend <- get_legend(legend)
# 
# #' Combine and write out.
# subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
#   subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
#   subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
#   subplots[[10]] + subplots[[11]] + subplots[[12]] +
#   plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))

```{r table-2}
cols_numeric <- c("r.squared", "adj.r.squared", "sigma", "statistic", "logLik", "AIC", "BIC")

linear_model_results %>%
  select(-fits) %>%
  arrange(factor(x_var, levels = order_vars)) %>%
  mutate(
    x_var = case_when(
      x_var == "rh_min" ~    "Min. Rel. Humidity",
      x_var == "par_total" ~ "DLI",
      x_var == "temp_mean" ~ "Mean Temp."
    ),
    y_var = case_when(
      y_var == "total_cover" ~ "Total cover",
      y_var == "count_mean" ~ "Gemmae count",
      y_var == "length_mean" ~ "Gemmae length"
    )
  ) %>%
  group_by(x_var) %>%
  gt() %>%
  cols_move_to_start(vars(y_var)) %>%
  cols_move(
    columns = vars(df.residual),
    after = vars(df)
  ) %>% 
  fmt_number(
    columns = colnames(.) %in% cols_numeric
  ) %>%
  fmt_number(
    columns = vars(p.value),
    decimals = 3
  ) %>%
  fmt_number(
    columns = vars(deviance),
    decimals = 0
  ) %>%
  cols_label(
    y_var = "",
    r.squared = "r-squared",
    adj.r.squared = "adj. r-squared",
    statistic = "t",
    p.value = "p",
    df.residual = "df residual"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 2"
  )

```{r fig-6, fig.width=7, fig.height=8}
#' Loop through all combination of dep and indep variables and their models,
#' and make plots for each.
var_grid <- make_var_grid (c("rh_min", "par_total", "temp_mean"))

subplots <- 
  pmap(var_grid, 
       quick_plot_with_stats, 
       plot_data = combined_monthly_morph, 
       model_data = linear_model_results)

#' Add axis titles
#' 
#' First, remove all default titles, then add formatted titles.
subplots <- map(subplots, ~ . + labs(x = "", y = ""))
subplots[[1]] <- subplots[[1]] + labs(y = expression("Total cover ("~cm^2~")"))
subplots[[4]] <- subplots[[4]] + labs(y = "Gemmae count")
subplots[[7]] <- subplots[[7]] + 
  labs(
    y = expression(paste("Gemmae len. (", mu, "m)", sep = "")),
    x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = ""))
  )
subplots[[8]] <- subplots[[8]] + labs(x = expression("Min. Rel. Humidity (%)"))
subplots[[9]] <- subplots[[9]] + labs(x = expression("Mean Temp. (°C)"))

#' Set amount of gap to close on left and top
left_close <- -15
top_close <- -10

#' Close gaps on left only: 2,3
subplots[2:3] <- subplots[2:3] %>%
  map(~ . + theme(plot.margin = margin(0,0,0,left_close)) 
  )

#' Close gaps on left side and top: 5,6,8,9
subplots[c(5,6,8,9)] <- subplots[c(5,6,8,9)] %>%
  map(~ . + theme(plot.margin = margin(top_close,0,0,left_close)) 
  )

#' Close gaps on top only: 4,7
subplots[c(4,7)] <- subplots[c(4,7)] %>%
  map(~ . + theme(plot.margin = margin(top_close,0,0,0)) 
  )

#' Add legend for middle bottom plot, combine, and write out.
subplots[[8]] <- subplots[[8]] + 
  theme(legend.position="bottom") + 
  labs(color = "Month", shape = "Year")

subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  subplots[[4]] + subplots[[5]] + subplots[[6]] + 
  subplots[[7]] + subplots[[8]] + subplots[[9]] + 
  plot_layout(ncol = 3)

#' Black and white version if needed.
# subplots <- 
#   map(subplots,
#       ~ . + scale_color_grey() 
#   )
# 
# subplots[[1]] + subplots[[2]] + subplots[[3]] + 
#   subplots[[4]] + subplots[[5]] + subplots[[6]] + 
#   subplots[[7]] + subplots[[8]] + subplots[[9]] + 
#   plot_layout(ncol = 3)
```

```{r fig-7, fig.width=7, fig.height=4}
#' Note there is one outlier with length 46 mm, 
#' then the rest all less than 19 mm. We will remove this 
#' before plotting and make a note in the figure caption.
gameto_size_no_outlier <-
  gameto_size %>%
  # Change sex to uppercase for nicer plotting
  mutate(sex = stringr::str_to_title(sex)) %>%
  # Remove one outlier
  filter(length < 40)

#' Make jitter plots overlayed with box plots. Make two
#' subplots: one for width and one for length, then combine.

# Use colorbrewer palette: 3 diverging, color-blind friendly colors
cols <- c("#1b9e77","#d95f02","#7570b3")

# Try to make sure dots stay in the same place, but doesn't work
set.seed(1000)

# Color version

@@

subplots[["length"]] <- 
  ggplot(gameto_size_no_outlier, aes(y = length, x = sex, color = sex)) +
  labs(
    subtitle = "a",
    y = "Length (mm)",
    x = "",
    color = "")

subplots[["width"]] <-
  ggplot(gameto_size_no_outlier, aes(y = width, x = sex, color = sex)) +
  labs(
    subtitle = "b",
    y = "Width (mm)",
    x = "")

subplots <- 
  map(subplots,
      ~ . + 
        geom_jitter(size = 3) +
        geom_boxplot(
          outlier.shape = NA,
          fill = NA,
          color = "dark grey",
          alpha = 0.7
        ) +
        scale_color_manual(values = cols) +
        theme(legend.position = "none")
  )

subplots[[1]] + subplots[[2]] + plot_layout(ncol = 2)

# Black and white version
# subplots <- 
#   map(subplots,
#       ~ . + 
#         scale_color_grey() +
#         theme(legend.position = "none")
#   )
